DOpElib: Differential Equations and Optimization Environment; A Goal Oriented Software Library for Solving PDEs and Optimization Problems

### 2012-11-09 Paper_May_2012/main.tex
Kurzer Kommentar zu reusability von Code im stationaeren/instationaeren Fall am Beispiel der CellEquation eingefuegt, zudem den von Thomas in der Mail angesprochenen Abschnitt abgeaendert.

The pressure and incompressibility terms should be 
treated fully implicitly when time-discretizing 
the Navier-Stokes equations.
For this reason, the implementation is split up
into several items to account for fully implicit 
terms and remaining terms for which we need 
old time-step and present time-step values. 
Hence, in the local cell equation, the parameters
\texttt{scale} and \texttt{scale\_ico}
are used to distinguish these terms:
\begin{lstlisting}
  void CellEquation(...)
   {
   // Usual loops over DoFs and quadrature points
     {
      local_cell_vector(i) +=  
          scale *// terms like (\nabla v^{n+1}, \nabla \phi) etc.
          + scale_ico *// terms like (-p^{n+1},\nabla\cdot \phi)
     }
   }
\end{lstlisting}
Analogous procedure is performed in the local 
cell matrix.

In addition to stationary problems, the implementation
of the time-derivative must be considered which is done
in 
\begin{lstlisting}
  void CellTimeEquation(...)
   {
      //implement (u^{n+1},\phi)
   }
\end{lstlisting}
% \begin{lstlisting}
%   void CellTimeEquation(...)
%    {
%      local_cell_vector(i) += 
%          scale * (u^{n+1},\phi) + etc. 
%    }
% \end{lstlisting}
and it is done in a straight-forward way:\marginpar{Mir gefaellt dieses Auseinanderziehen nicht so wirklich...ist dieser zweite Abschnitt heir notwendig?}
\begin{lstlisting}
  for (/*all quadrature points per cell q*/)
   {
     Tensor<1,2> u = ...;

     for (/*all dofs per cell i*/)
      {
        Tensor<1,2> phi_i_u = ...;
        local_cell_vector(i) +=  scale * (u * phi_i_u) * 
                                  state_fe_values.JxW(q_point);    
      }
   }
\end{lstlisting}
Here, the advantage is that \dope{} orders the terms in the time-stepping
problems
such that only $(u^{n+1},\phi)$ must be implemented.

\begin{remark}[Nonlinear terms in the time derivative]
However, if someone 
needs explicit input of the old time-step term, then he can use 
\begin{lstlisting}
  void CellTimeEquationExplicit (...)
   {
     // implement (u^{n+1},\phi) and (u^n,\phi)  
   }
\end{lstlisting}
\end{remark}

>>>>

The pressure and incompressibility terms should be 
treated fully implicitly when time-discretizing 
the Navier-Stokes equations.
For this reason, the implementation is split up
into several items to account for fully implicit 
terms and remaining terms for which we need 
old time-step and present time-step values. 
Hence, in the local cell equation, the parameters
\texttt{scale} and \texttt{scale\_ico}
are used to distinguish these terms:
\begin{lstlisting}
  void CellEquation(...)
   {
   // Usual loops over DoFs and quadrature points
     {
      local_cell_vector(i) +=  
          scale *// terms like (\nabla v^{n+1}, \nabla \phi) etc.
          + scale_ico *// terms like (-p^{n+1},\nabla\cdot \phi)
     }
   }
\end{lstlisting}
Analogous procedure is performed in the local 
cell matrix.
\begin{remark}
Note that in the stationary case \dope{} automatically sets $ \texttt{scale\_ico} = \texttt{scale}$, so that in fact the user has to implement the \texttt{CellEquation} only once and can henceforth reuse it in all his stationary an instationary problems. This transfers of course also to other methods like \texttt{CellMatrix}.
\end{remark}

In addition to stationary problems, the implementation
of the time-derivative, i.e. the term $ (v^{n+1},\phi)$, must be considered which is done
in the method \texttt{CellTimeEquation} in a straight-forward way. After the initialization of $v$ and the shape function $\phi_i$ and the usual loops over the quadrature points and  degrees of freedom of the cell we fill the local cell vector by
\begin{lstlisting}
        local_cell_vector(i) +=  scale * (v * phi_i) * 
                                  state_fe_values.JxW(q_point);    
\end{lstlisting}
Here, the advantage is that \dope{} orders the terms in the time-stepping
problems
such that only $(v^{n+1},\phi)$ must be implemented.

\begin{remark}[Nonlinear terms in the time derivative]
However, if someone 
needs explicit input of the old time-step term, then he can use 
\begin{lstlisting}
  void CellTimeEquationExplicit (...)
   {
     // implement (v^{n+1},\phi) and (v^n,\phi)  
   }
\end{lstlisting}
\end{remark}
